---
title: "Variation in infectious period"
output: html_document
---

First, let's define a function that draws $r-R$ curve given shape and scale parameters of latent and infectious periods.

```{r}
genfun <- function(lat.shape=2,
                   lat.scale=1,
                   inf.shape=1,
                   inf.scale=1,
                   n=10000,
                   r=seq(0, 2, by=0.1),
                   add=FALSE,
                   ...
                   ) {
    set.seed(101)
    lat <- rgamma(n, lat.shape, lat.scale)
    inf <- rgamma(n, inf.shape, inf.scale)
    inf <- sample(inf, n, replace=TRUE, prob=inf)

    gen <- lat + runif(n, 0, inf)
    
    if (!add) {
        plot(sapply(r, function(x) 1/mean(exp(-x*gen))), type="l", ...)
    } else{
        lines(sapply(r, function(x) 1/mean(exp(-x*gen))), ...)
    }
    
    invisible(gen)
}
```

As we decrease variation in infectious period, while keeping mean fixed, we see increase in $R$ given $r$.

```{r}
genfun()
genfun(inf.shape=2, inf.scale=2, add=TRUE, lty=2)
genfun(inf.shape=3, inf.scale=3, add=TRUE, lty=3)
```

What happens if contacts are concentrated in the early-phase of infectious period?
Here's another function:

```{r}
genfun2 <- function(lat.shape=2,
                   lat.scale=1,
                   inf.shape=1,
                   inf.scale=1,
                   n=10000,
                   r=seq(0, 2, by=0.1),
                   add=FALSE,
                   ...
                   ) {
    set.seed(101)
    lat <- rgamma(n, lat.shape, lat.scale)
    inf <- rgamma(n, inf.shape, inf.scale)
    inf <- sample(inf, n, replace=TRUE, prob=inf)

    gen <- lat + rbeta(n, 1, 9) * inf
    
    if (!add) {
        plot(sapply(r, function(x) 1/mean(exp(-x*gen))), type="l", ...)
    } else{
        lines(sapply(r, function(x) 1/mean(exp(-x*gen))), ...)
    }
    
    invisible(gen)
}
```

As we decrease variation in infectious period, while keeping mean fixed, we see increase in $R$ given $r$.

```{r}
genfun2()
genfun2(inf.shape=2, inf.scale=2, add=TRUE, lty=2)
genfun2(inf.shape=3, inf.scale=3, add=TRUE, lty=3)
```

